---
title: "2-DischargeEmpirical"
author: "Anıl Axel Tellbüscher"
date: '2022-08-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# General
A constraint that is limiting the amount of a substance in an aquaculture system beside its solubility is the water exchange rate. After approximately three times the hydraulic retention time (HRT), the substance inflow concentration in the tank equals the outflow concentratino according to the CFSTR model. Even though this is an idealised model, it has to be considered that the outflow concentration can not be greater as the inflow concentration, because no generation of the substances of interest is taking place. Instead, the real outflow concentration is expected to be lower than the ideal concentration due to microbial uptake and assimilation by fish.

The required data to calculate theoretical asymptotic concentrations of a substance are
- all mass inputs of the substance, that are
  - feed
  - makeup water
  - auxiliary substances
- HRT, being calculated using
  - total system volume
  - water exchange rate
  
  
# Preparations

## Load packages
```{r}
library(XLConnect)
library(tidyverse)
```



## Load data
```{r Load workbook}
# Load workbook
masterfile_nutrients <- loadWorkbook(filename = "../2 - Nutrient Solubility/data/Aquaponics - Masterfile Nutrients.xlsx")
```



```{r Load worksheets}
# Samples to be considered
samples <- readWorksheet(masterfile_nutrients, 'waterAVERAGE', endCol = 2)

# Rearing data
rearing <- readWorksheet(masterfile_nutrients, 'rearing', startRow = 2)

# Feed data
feed <- readWorksheet(masterfile_nutrients, 'feedIN', startRow = 2, endCol = 30)
  
# Makeup water
water <- readWorksheet(masterfile_nutrients, 'waterIN', startRow = 2, endCol = 26)
```



## Tidy data
Initially, only the observations that were already pre-selected during the analysis of nutrient concentrations etc. should be used. Thus, the total observations are united with the 'samples' dataset, containing only the identifiers of the pre-selected observations.

```{r}
procdata <- samples %>% 
  left_join(rearing) %>% 
  left_join(feed) %>% 
  left_join(water)
```



# Calculations

## Hydraulic Retention Time

```{r}
procdata <- procdata %>% 
  mutate(
    hrt = Vtot_m3 / (Vtot_m3 * exchRate)
  )
```



## Feed input

```{r}
procdata <- procdata %>% 
  mutate(
    feed_kgm3 = feed_kgd / (Vtot_m3 * exchRate), # Input per m3
  )
```

