---
title: "Aquaponics versus Hydroponics"
subtitle: "Comparison of nutrient concentrations recommended for hydroponics with their maximum solubilities under aquaponic conditions"
author: "Anıl A. Tellbüscher"
date: '2022-07-07'
output: html_document
---

```{r}
library(XLConnect)
library(tidyverse)
```



# General
Conditions in aquaponic systems differ from those in hydroponics. The main differences with regards to physicochemical conditions are

- pH
- DOM


The question arising is now whether the recommended elemental concentrations for hydroponic nutrient solutions can actually be reached in aquaponic systems from a chemical viewpoint or whether precipitation occurs due to the higher pH in the system.







# 1. Hydroponic data
Nutrient solution composition data presented by Resh (Hydroponic Food Production) was extracted from the book and used for the following calculations.



## 1.1 Load data
```{r Resh Workbook loading}
# Load Workbook
resh_hydroponic_plant_production <- loadWorkbook(filename = "/Users/anil/Library/CloudStorage/OneDrive-JihočeskáuniverzitavČeskýchBudějovicích/General/Resh - Hydroponic Plant Production.xlsx")
```



```{r Resh Workbook reading}
# Read Worksheet
# Unit: mg/L
resh_data <- readWorksheet(resh_hydroponic_plant_production, sheet = "nutrient solutions", region = c('A', 'C2:Q50'))
```



```{r Molar weight data}
molarWeight <- read.csv2('data/molarWeight.csv')
```



## 1.2 Data wrangling
```{r}
resh_data <- data.frame(resh_data[[2]], row.names = resh_data[[1]][2:49,1])

resh_data <- resh_data %>% 
  rename('P' = 'PO4_P_mgL', 'K' = 'K_mgL', 'Ca' = 'Ca_mgL', 'Mg' = 'Mg_mgL', 'S' = 'SO4_S_mgL', 'B' = 'B_mgL', 'Fe' = 'Fe_mgL', 'Mn' = 'Mn_mgL', 'Cu' = 'Cu_mgL', 'Zn' = 'Zn_mgL', 'Na' = 'Na_mgL', 'NH4' = 'NH4_N_mgL', 'Mo' = 'Mo_mgL', 'NO3' = 'NO3_N_mgL', 'Cl' = 'Cl_mgL') 

print(resh_data)
```



## 1.3 Create dataset with target concentrations in hydroponics
The nutrient data is all given in mg/L. These mass concentrations are not suitable for the correction by pH. Thus, data has to be converted into molar concentrations for further calculations.



```{r}
targetConc <- resh_data %>% 
  
  # Calculate total N (TIN) by summing up NH4 and NO3
  mutate(
    N = NH4 + NO3
  ) %>% 
  select(-c('NH4', 'NO3')) %>% 
  
  # Create longtable
  pivot_longer(
    everything(),
    names_to = 'nutrient',
    values_to = 'conc_mgL'
    ) %>% 
  
  drop_na() %>%   # Remove missing values
  
  #group_by(nutrient) %>% 
  #summarise(mean_mgL = mean(concentration)) %>% 
  
  left_join(molarWeight, by = c('nutrient' = 'element')) %>% 
  
  mutate(
    g_per_mol = as.numeric(g_per_mol),
    mol_per_L = conc_mgL * 1e-3 / g_per_mol # Conversion into mol/L
    ) %>% 
  
  arrange(desc(mol_per_L)) %>% 
  
  print()
```  



## 1.4 Recalculation considering environmental aspects
Due to the fact that conditions in hydroponic systems differ from those observed in aquaculture, some assumptions can be made such as the pH being maintained at acidic levels. Furthermore, it can be assumed that the $CO_{2}$ saturation of the water is likely comparably high because no gas stripping technology is applied and in some greenhouses the atmospheric $CO_{2}$ concentration is even artificially increased.
The pH and $p(CO_{2})$ values utilized for the following calculations are thus presentedin the following table.



|Variable   |Assumption|
|-----------|----------|
|pH         |6.0       |
|$p(CO_{2})$|0.54 atm  |



```{r Assumptions, include=FALSE}
pH <- 6
pCO2_atm <- 0.054
```



The correction of the concentrations for pH can eventually be done using self-written speciation functions.

```{r Load species correction functions}
# Phosphate
source('R/speciation - c_T-PO4.R')

# Hydroxide
source('R/speciation - OH.R')

# Carbonate
source('R/speciation - pCO2-CO3.R')

# Sulfate
source('R/speciation - c_T-SO4.R')
```



Eventually, the functions can be applied to the dataset.
**The extra idented rows are a workaround to make *pivot_wider()* work.**

```{r}
procdata <- targetConc %>%
  
  select(c(nutrient, mol_per_L)) %>% 
  
      group_by(nutrient) %>% # Workaround pivot_wider()
      mutate(row = row_number()) %>% # Workaround pivot_wider()
  
  pivot_wider(names_from = nutrient, values_from = mol_per_L) %>% 
  
      select(-row) %>%  # Workaround pivot_wider()

  
  # Calculate species
  mutate(
    species_PO4 = calc_PO4(c_tot = P, pH = pH, form_P = 'po4-p', unit = 'mol'),
    
    species_OH = calc_OH(pH = pH),
    
    species_CO3 = calc_CO3(pH = pH, pCO2 = pCO2_atm)
  ) %>% 
  
  print()
```



```{r}  
procdata %>% 
  pivot_longer(
    cols = everything(),
    names_to = 'nutrient',
    values_to = 'conc_molL'
  ) %>% 
  
  drop_na() %>% 
  
  ggplot(aes(y = conc_molL)) + 
  geom_boxplot() + 
  facet_wrap(facets = vars(nutrient), scales = 'free')
```



The next step after the pH correction is to calculate theoretical lowest ('worst case') solubilities of all compounds.

```{r}
# Import solubility data
solubilityData <- read.csv('data/solubilityData_for4-NutrientComparison.csv')

#
procdata %>% 
  pivot_wider(
    names_from = 'nutrient',
    values_from = 'conc_molL'
  ) %>% 
  
  # Now same as in 3-NutrinetSolubility line 182

```

