---
title: "Aquaponics versus Hydroponics"
subtitle: "Comparison of nutrient concentrations recommended for hydroponics with their maximum solubilities under aquaponic conditions"
author: "Anıl A. Tellbüscher"
date: '2022-07-07'
output: html_document
---

```{r}
library(XLConnect)
library(tidyverse)
```



# General
Conditions in aquaponic systems differ from those in hydroponics. The main differences with regards to physicochemical conditions are

- pH
- DOM


The question arising is now whether the recommended elemental concentrations for hydroponic nutrient solutions can actually be reached in aquaponic systems from a chemical viewpoint or whether precipitation occurs due to the higher pH in the system.







# 1. Preparations
Nutrient solution composition data presented by Resh (Hydroponic Food Production) was extracted from the book and used for the following calculations.



## 1.1 Load data
```{r Resh Workbook loading}
# Load Workbook
resh_hydroponic_plant_production <- loadWorkbook(filename = "/Users/anil/Library/CloudStorage/OneDrive-JihočeskáuniverzitavČeskýchBudějovicích/General/Resh - Hydroponic Plant Production.xlsx")
```



```{r Resh Workbook reading}
# Read Worksheet
# Unit: mg/L
resh_data <- readWorksheet(resh_hydroponic_plant_production, sheet = "nutrient solutions", region = c('A', 'C2:Q50'))
```



```{r Molar weight data}
molarWeight <- read.csv2('data/molarWeight.csv')
```



## 1.2 Data wrangling
```{r}
resh_data <- data.frame(resh_data[[2]], row.names = resh_data[[1]][2:49,1])

resh_data <- resh_data[7,]

resh_data <- resh_data %>% 
  rename('P' = 'PO4_P_mgL', 'K' = 'K_mgL', 'Ca' = 'Ca_mgL', 'Mg' = 'Mg_mgL', 'S' = 'SO4_S_mgL', 'B' = 'B_mgL', 'Fe' = 'Fe_mgL', 'Mn' = 'Mn_mgL', 'Cu' = 'Cu_mgL', 'Zn' = 'Zn_mgL', 'Na' = 'Na_mgL', 'NH4' = 'NH4_N_mgL', 'Mo' = 'Mo_mgL', 'NO3' = 'NO3_N_mgL', 'Cl' = 'Cl_mgL') 

print(resh_data)
```



## 1.3 Create dataset with target concentrations in hydroponics
The nutrient data is all given in mg/L. These mass concentrations are not suitable for the correction by pH. Thus, data has to be converted into molar concentrations for further calculations.



```{r}
targetConc <- resh_data %>% 
  
  # Calculate total N (TIN) by summing up NH4 and NO3
  mutate(
    N = NH4 + NO3
  ) %>% 
  select(-c('NH4', 'NO3')) %>% 
  
  # Create longtable
  pivot_longer(
    everything(),
    names_to = 'nutrient',
    values_to = 'conc_mgL'
    ) %>% 
  
  drop_na() %>%   # Remove missing values
  
  #group_by(nutrient) %>% 
  #summarise(mean_mgL = mean(concentration)) %>% 
  
  left_join(molarWeight, by = c('nutrient' = 'element')) %>% 
  
  mutate(
    g_per_mol = as.numeric(g_per_mol),
    mol_per_L = conc_mgL * 1e-3 / g_per_mol # Conversion into mol/L
    ) %>% 
  
  arrange(desc(mol_per_L)) %>% 
  
  print()
```  



## 1.4 Recalculation considering environmental aspects
Due to the fact that conditions in hydroponic systems differ from those observed in aquaculture, some assumptions can be made such as the pH being maintained at acidic levels. Furthermore, it can be assumed that the $CO_{2}$ saturation of the water is likely comparably high because no gas stripping technology is applied and in some greenhouses the atmospheric $CO_{2}$ concentration is even artificially increased.
The pH and $p(CO_{2})$ values utilized for the following calculations are thus presentedin the following table.



|Variable   |Assumption|
|-----------|----------|
|pH         |6.0       |
|$p(CO_{2})$|0.54 atm  |



```{r Assumptions, include=FALSE}
# Values in hydroponics
pH <- 6
pCO2_atm <- 0.054

# Values in aquaculture/aquaponics
pH2 <- 7.5
pCO2_atm2 <- 0.054
```



The correction of the concentrations for pH can eventually be done using self-written speciation functions.

```{r Load species correction functions}
# Phosphate
source('R/speciation - c_T-PO4.R')

# Hydroxide
source('R/speciation - OH.R')

# Carbonate
source('R/speciation - pCO2-CO3.R')

# Sulfate
source('R/speciation - c_T-SO4.R')
```



Eventually, the functions can be applied to the dataset.
**The extra idented rows are a workaround to make *pivot_wider()* work.**

```{r}
procdata <- targetConc %>%
  
  select(c(nutrient, mol_per_L)) %>% 
  
      group_by(nutrient) %>% # Workaround pivot_wider()
      mutate(row = row_number()) %>% # Workaround pivot_wider()
  
  pivot_wider(names_from = nutrient, values_from = mol_per_L) %>% 
  
      select(-row) %>%  # Workaround pivot_wider()

  
  # Calculate species
  mutate(
    species_PO4_molL = calc_PO4(c_tot = P, pH = pH, form_P = 'po4-p', unit = 'mol'),
    
    species_OH_molL = calc_OH(pH = pH),
    
    species_CO3_molL = calc_CO3(pH = pH, pCO2 = pCO2_atm),
    
    species_SO4_molL = calc_SO4(c_tot = S, pH = pH, form_S = 'SO4-S', unit = 'mol')
  ) %>% 
  
  print()
```



```{r}  
procdata %>% 
  
  select(
    -c(
      P
      ,S
      ,species_OH_molL
      ,species_CO3_molL
    )
  ) %>% 
  
  pivot_longer(
    cols = everything(),
    names_to = 'nutrient',
    values_to = 'conc_molL'
  ) %>% 
  
  drop_na() %>% 
  
  ggplot(aes(y = conc_molL)) + 
  geom_boxplot() + 
  facet_wrap(facets = vars(nutrient), scales = 'free')
```



The next step after the pH correction is to calculate theoretical lowest ('worst case') solubilities of all compounds ('X_in_XY' denoting for the theoretical maximum concentration of a compound under the given conditions).
Furthermore, a logical test is conducted, assessing whether the recommended concentration is larger than the theoretical solubility. Also, the difference between the recommendation and the theoretical solubility and their proportions are calculated.

```{r}
# Import solubility data
solubility_data <- read.csv('data/solubilityData_for4-NutrientComparison.csv')

#
plotdata <- procdata %>% 

mutate(
    # CaCO3
    Ca_in_CaCO3 = solubility_data$Ks0[which(solubility_data$chemFormula == "CaCO3",)] / species_CO3_molL,
    Ca_found_greater_CaCO3 = Ca > Ca_in_CaCO3,
    CaCO3_diff = Ca - Ca_in_CaCO3,
    CaCO3_prop = Ca / Ca_in_CaCO3,
    
    # Ca(OH)2
    Ca_in_CaOH2 = solubility_data$Ks0[which(solubility_data$chemFormula == "Ca(OH)2",)] / species_OH_molL^2,
    Ca_found_greater_CaOH2 = Ca > Ca_in_CaOH2,
    CaOH2_diff = Ca - Ca_in_CaOH2,
    CaOH2_prop = Ca / Ca_in_CaOH2,
    
    # Ca3PO42
    Ca_in_Ca3PO42 = (solubility_data$Ks0[which(solubility_data$chemFormula == "Ca3(PO4)2",)] / species_PO4_molL^2)^(1/3),
    Ca_found_greater_Ca3PO42 = Ca > Ca_in_Ca3PO42,
    Ca3PO42_diff = Ca - Ca_in_Ca3PO42,
    Ca3PO42_prop = Ca / Ca_in_Ca3PO42,

    
    # CaSO4
    Ca_in_CaSO4 = solubility_data$Ks0[which(solubility_data$chemFormula == "CaSO4",)] / species_SO4_molL,
    Ca_found_greater_CaSO4 = Ca > Ca_in_CaSO4,
    CaSO4_diff = Ca - Ca_in_CaSO4,
    CaSO4_prop = Ca / Ca_in_CaSO4,
    
    # MgCO3
    Mg_in_MgCO3 = solubility_data$Ks0[which(solubility_data$chemFormula == "MgCO3",)] / species_CO3_molL,
    Mg_found_greater_MgCO3 = Mg > Mg_in_MgCO3,
    MgCO3_diff = Mg - Mg_in_MgCO3,
    MgCO3_prop = Mg / Mg_in_MgCO3,
    
    # Mg(OH)2
    Mg_in_MgOH2 = solubility_data$Ks0[which(solubility_data$chemFormula == "Mg(OH)2",)] / species_OH_molL^2,
    Mg_found_greater_MgOH2 = Mg > Mg_in_MgOH2,
    MgOH2_diff = Mg - Mg_in_MgOH2,
    MgOH2_prop = Mg / Mg_in_MgOH2,

    
    # Mg3PO42
    Mg_in_Mg3PO42 = (solubility_data$Ks0[which(solubility_data$chemFormula == "Mg3(PO4)2",)] / species_PO4_molL^2)^(1/3),
    Mg_found_greater_Mg3PO42 = Mg > Mg_in_Mg3PO42,
    Mg3PO42_diff = Mg - Mg_in_Mg3PO42,
    Mg3PO42_prop = Mg / Mg_in_Mg3PO42,

    # FePO4
    Fe_in_FePO4 = solubility_data$Ks0[which(solubility_data$chemFormula == "FePO4 x 2 H2O",)] / species_PO4_molL,
    Fe_found_greater_FePO4 = Fe > Fe_in_FePO4,
    FePO4_diff = Fe - Fe_in_FePO4,
    FePO4_prop = Fe / Fe_in_FePO4,
    
    # MnCO3
    Mn_in_MnCO3 = solubility_data$Ks0[which(solubility_data$chemFormula == "MnCO3",)] / species_CO3_molL,
    Mn_found_greater_MnCO3 = Mn > Mn_in_MnCO3,
    MnCO3_diff = Mn - Mn_in_MnCO3,
    MnCO3_prop = Mn / Mn_in_MnCO3,
    
    # Cu3PO42
    Cu_in_Cu3PO42 = (solubility_data$Ks0[which(solubility_data$chemFormula == "Cu3(PO4)2",)] / species_PO4_molL^2)^(1/3),
    Cu_found_greater_Cu3PO42 = Cu > Cu_in_Cu3PO42,
    Cu3PO42_diff = Cu - Cu_in_Cu3PO42,
    Cu3PO42_prop = Cu / Cu_in_Cu3PO42,
    
    # ZnCO3
    Zn_in_ZnCO3 = solubility_data$Ks0[which(solubility_data$chemFormula == "ZnCO3",)] / species_CO3_molL,
    Zn_found_greater_ZnCO3 = Zn > Zn_in_ZnCO3,
    ZnCO3_diff = Zn - Zn_in_ZnCO3,
    ZnCO3_prop = Zn / Zn_in_ZnCO3,
    
    # Zn(OH)2
    Zn_in_ZnOH2 = solubility_data$Ks0[which(solubility_data$chemFormula == "Zn(OH)2",)] / species_OH_molL^2,
    Zn_found_greater_ZnOH2 = Zn > Zn_in_ZnOH2,
    ZnOH2_diff = Zn - Zn_in_ZnOH2,
    ZnOH2_prop = Zn / Zn_in_ZnOH2
  )
```





# 2. Comparison between recommendation and theoretical maximum concentration
In the following, the dataset is reduced to the calculated proportions between the concentrations that are recommended and the concentrations that are resulting from the theoretical solubility calculations

```{r}
plotdata <- plotdata %>% 
  
  select(ends_with('_prop')) %>% 
  
  pivot_longer(
    cols = everything(),
    names_to = 'substance',
    values_to = 'proportion'
  ) %>% 
  
  mutate(
    substance = sub('_prop', '', substance),
    pH = pH,
    pCO2 = pCO2_atm
  ) %>% 
  
  filter(
    proportion != Inf
    ) %>% 
  
  print()
```




```{r}
plotdata %>% 
  
  mutate(
    substance = as.factor(substance),
    treshold_label = if_else(proportion > 1, true = '2', false = '1')
    ) %>% 
  
  ggplot(aes(
    x = substance,
    y = proportion,
    fill = treshold_label
    )) + 
    geom_boxplot() + 
    geom_hline(yintercept = 1, linetype = 'dashed', color = 'red', alpha = 0.75) +
    scale_y_log10() + 
    #facet_wrap(facets = vars(substance), scales = 'free') + 
    labs(
      x = 'Recommendation compared with theoretical solubility',
      y = 'log(proportion)'
    ) + 
    theme_minimal() + 
    theme(
      axis.text.x = element_text(angle = 45),
      legend.position = 'none'
      )
```



š

# 3. Comparison between different environmental conditions






```{r Calculate species concentrations under secondary conditions for comparison}
procdata2 <- targetConc %>%
  
  select(c(nutrient, mol_per_L)) %>% 
  
      group_by(nutrient) %>% # Workaround pivot_wider()
      mutate(row = row_number()) %>% # Workaround pivot_wider()
  
  pivot_wider(names_from = nutrient, values_from = mol_per_L) %>% 
  
      select(-row) %>%  # Workaround pivot_wider()

  
  # Calculate species
  mutate(
    species_PO4_molL = calc_PO4(c_tot = P, pH = pH2, form_P = 'po4-p', unit = 'mol'),
    
    species_OH_molL = calc_OH(pH = pH2),
    
    species_CO3_molL = calc_CO3(pH = pH2, pCO2 = pCO2_atm2),
    
    species_SO4_molL = calc_SO4(c_tot = S, pH = pH2, form_S = 'SO4-S', unit = 'mol')
  ) %>% 
  
  print()
```




```{r Calculate resulting solubility data under secondary conditions for comparison}
plotdata2 <- procdata2 %>% 

mutate(
    # CaCO3
    Ca_in_CaCO3 = solubility_data$Ks0[which(solubility_data$chemFormula == "CaCO3",)] / species_CO3_molL,
    Ca_found_greater_CaCO3 = Ca > Ca_in_CaCO3,
    CaCO3_diff = Ca - Ca_in_CaCO3,
    CaCO3_prop = Ca / Ca_in_CaCO3,
    
    # Ca(OH)2
    Ca_in_CaOH2 = solubility_data$Ks0[which(solubility_data$chemFormula == "Ca(OH)2",)] / species_OH_molL^2,
    Ca_found_greater_CaOH2 = Ca > Ca_in_CaOH2,
    CaOH2_diff = Ca - Ca_in_CaOH2,
    CaOH2_prop = Ca / Ca_in_CaOH2,
    
    # Ca3PO42
    Ca_in_Ca3PO42 = (solubility_data$Ks0[which(solubility_data$chemFormula == "Ca3(PO4)2",)] / species_PO4_molL^2)^(1/3),
    Ca_found_greater_Ca3PO42 = Ca > Ca_in_Ca3PO42,
    Ca3PO42_diff = Ca - Ca_in_Ca3PO42,
    Ca3PO42_prop = Ca / Ca_in_Ca3PO42,

    
    # CaSO4
    Ca_in_CaSO4 = solubility_data$Ks0[which(solubility_data$chemFormula == "CaSO4",)] / species_SO4_molL,
    Ca_found_greater_CaSO4 = Ca > Ca_in_CaSO4,
    CaSO4_diff = Ca - Ca_in_CaSO4,
    CaSO4_prop = Ca / Ca_in_CaSO4,
    
    # MgCO3
    Mg_in_MgCO3 = solubility_data$Ks0[which(solubility_data$chemFormula == "MgCO3",)] / species_CO3_molL,
    Mg_found_greater_MgCO3 = Mg > Mg_in_MgCO3,
    MgCO3_diff = Mg - Mg_in_MgCO3,
    MgCO3_prop = Mg / Mg_in_MgCO3,
    
    # Mg(OH)2
    Mg_in_MgOH2 = solubility_data$Ks0[which(solubility_data$chemFormula == "Mg(OH)2",)] / species_OH_molL^2,
    Mg_found_greater_MgOH2 = Mg > Mg_in_MgOH2,
    MgOH2_diff = Mg - Mg_in_MgOH2,
    MgOH2_prop = Mg / Mg_in_MgOH2,

    
    # Mg3PO42
    Mg_in_Mg3PO42 = (solubility_data$Ks0[which(solubility_data$chemFormula == "Mg3(PO4)2",)] / species_PO4_molL^2)^(1/3),
    Mg_found_greater_Mg3PO42 = Mg > Mg_in_Mg3PO42,
    Mg3PO42_diff = Mg - Mg_in_Mg3PO42,
    Mg3PO42_prop = Mg / Mg_in_Mg3PO42,

    # FePO4
    Fe_in_FePO4 = solubility_data$Ks0[which(solubility_data$chemFormula == "FePO4 x 2 H2O",)] / species_PO4_molL,
    Fe_found_greater_FePO4 = Fe > Fe_in_FePO4,
    FePO4_diff = Fe - Fe_in_FePO4,
    FePO4_prop = Fe / Fe_in_FePO4,
    
    # MnCO3
    Mn_in_MnCO3 = solubility_data$Ks0[which(solubility_data$chemFormula == "MnCO3",)] / species_CO3_molL,
    Mn_found_greater_MnCO3 = Mn > Mn_in_MnCO3,
    MnCO3_diff = Mn - Mn_in_MnCO3,
    MnCO3_prop = Mn / Mn_in_MnCO3,
    
    # Cu3PO42
    Cu_in_Cu3PO42 = (solubility_data$Ks0[which(solubility_data$chemFormula == "Cu3(PO4)2",)] / species_PO4_molL^2)^(1/3),
    Cu_found_greater_Cu3PO42 = Cu > Cu_in_Cu3PO42,
    Cu3PO42_diff = Cu - Cu_in_Cu3PO42,
    Cu3PO42_prop = Cu / Cu_in_Cu3PO42,
    
    # ZnCO3
    Zn_in_ZnCO3 = solubility_data$Ks0[which(solubility_data$chemFormula == "ZnCO3",)] / species_CO3_molL,
    Zn_found_greater_ZnCO3 = Zn > Zn_in_ZnCO3,
    ZnCO3_diff = Zn - Zn_in_ZnCO3,
    ZnCO3_prop = Zn / Zn_in_ZnCO3,
    
    # Zn(OH)2
    Zn_in_ZnOH2 = solubility_data$Ks0[which(solubility_data$chemFormula == "Zn(OH)2",)] / species_OH_molL^2,
    Zn_found_greater_ZnOH2 = Zn > Zn_in_ZnOH2,
    ZnOH2_diff = Zn - Zn_in_ZnOH2,
    ZnOH2_prop = Zn / Zn_in_ZnOH2
  )
```





```{r Convert comparison data into longtable and add pH and pCO2 identifier}
plotdata2 <- plotdata2 %>% 
  
  select(ends_with('_prop')) %>% 
  
  pivot_longer(
    cols = everything(),
    names_to = 'substance',
    values_to = 'proportion'
  ) %>% 
  
  mutate(
    substance = sub('_prop', '', substance),
    pH = pH2,
    pCO2 = pCO2_atm2
  ) %>% 
  
  filter(
    proportion != Inf
    ) %>% 
  
  print()
```





```{r Plot data for comparison}
plotdata %>% 
  
  bind_rows(plotdata2) %>% 
  
  mutate(
    substance = as.factor(substance),
    pH = as.factor(pH),
    pCO2 = as.factor(pCO2),
    treshold_label = if_else(proportion > 1, true = '2', false = '1')
    ) %>% 
  
  filter(
    !grepl('OH', substance) &
      !grepl('SO4', substance)
  ) %>% 
  
  ggplot(aes(
    x = substance,
    y = proportion,
    shape = pH,
    color = pCO2
    )) + 
    geom_point() + 

    geom_hline(yintercept = 1, linetype = 'dashed', color = 'red', alpha = 0.75) +
  
    scale_y_log10() + 
  
    facet_wrap(facets = vars(substance), scales = 'free') + 
  
    labs(
      x = 'Recommendation compared with theoretical solubility',
      y = 'log(proportion)'
    ) + 
    theme_minimal() + 
    theme(
      axis.text.x = element_blank()
      )
```